---
title: "chem explore"
theme: sandstone
format:
  html:
    css: styles.css
    grid:
      sidebar-width: 0px
      margin-width: 0px
      body-width: 2500px
---

```{r, library}
#| include: false

pacman::p_load(shiny,
               bslib,
               ggplot2,
               ggtext,
               shinythemes,
               thematic,
               tidyverse,
               httr,
               leaflet,
               leafem,
               terra,
               htmltools,
               glue,
               cowplot,
               leaflegend,
               smoothr,
               leaflet.providers,
               janitor,
               openxlsx,
               bsicons)

if (!requireNamespace("leaflet.extras", quietly = TRUE)) {
  remotes::install_github("trafficonese/leaflet.extras")
}
library(leaflet.extras)

knitr::read_chunk('addGroupedLayersControl.R')
```

```{r, import}
#| include: false
# import layers

## river extent
exploits_drainage <- vect("data/spatial data/exploits_watershed.shp") |>  
    project("EPSG:4326 - WGS 84")
  
## all rivers
exploits_river <- vect("data/spatial data/exploits_river_field_data.shp") |>  
  project("EPSG:4326 - WGS 84")
  
## lakes
lakes <- vect("data/spatial data/Exploits_lakes.shp") |>  
  project("EPSG:4326 - WGS 84")
  
## shape of NL
NL <- vect("data/spatial data/NL_boundaries.shp") |>  
  project("EPSG:4326 - WGS 84")  

## areas blocked by CABD 
file_barriers <- paste0("data/spatial data/exploits_barriers.shp")
barriers <- vect(file_barriers)

file_inaccess <- paste0("data/spatial data/exploits_inaccess.shp")
inaccess <- vect(file_inaccess)

# field data:
## invert
invert_field <- read.xlsx("data/Fielddata_2024_2025.xlsx", sheet = "Invertebrate and periphyton") |> 
  clean_names()

invert_field$y <- as.numeric(invert_field$y)
invert_field$x <- as.numeric(invert_field$x)

## parr and fry
parrfry_field <- read.xlsx("data/Fielddata_2024_2025.xlsx", sheet = "Juvenile fish 2025")|> 
  clean_names()


parrfry_field$y <- as.numeric(parrfry_field$y)
parrfry_field$x <- as.numeric(parrfry_field$x)

## smolts


## gravel
## data from river reaches sampled
gravel_field <- vect("data/spatial data/FieldSites.shp") |>  # to be updated with 2025 sites
  project("EPSG:4326 - WGS 84")

## water chem field data
chem <- read.xlsx("data/Water_Quality_data_explore.xlsx", sheet = "Data")
   
# import arrow
arrow <- "https://cdn.pixabay.com/photo/2013/07/12/17/54/arrow-152596_960_720.png"
```

```{r, basemap}
#| include: false

# base map gets specific layers added in each tab to it to make unique maps but with same base

exploits_map <- leaflet(width = "100vw", height = "72vh", options = leafletOptions(zoomSnap = 0.25)) |> 
  addProviderTiles('Esri.WorldImagery', group = "Satellite") |> # add base map options
  addProviderTiles('Esri.WorldTopoMap', group = "Topographic") |> # adds base map
  addProviderTiles('Esri.WorldGrayCanvas', group = "White & Grey") |> 

# open map to this screen
  setView(-56.65499, 48.70449, zoom = 8.5) |> 
  
#can recenter to that location by clicking button
  addHomeButton(ext = c(-58.2,48.1, -55.45,49.3), position = "topleft", group = "Recenter") |> 

# create layers to add vectors to
  addMapPane("NL", zIndex = 400) |> 
  addMapPane("drainage", zIndex = 410) |> 
  addMapPane("river", zIndex = 415) |> 
  addMapPane("lakes", zIndex = 420) |>
      
  # addPolygons(data = NL, color = "#A1929B", weight = 1.5, fillColor = "#A1929B", fillOpacity = 0.25, options = pathOptions(pane = "NL")) |>
      
# add polygons and lines
## exploits drainage area
  addPolygons(data = exploits_drainage, color = "#32B1DB", weight = 1.25, fillColor = "#236D9E", opacity = 0.4, fillOpacity = 0.2, options = pathOptions(pane = "drainage")) |>
  
## rivers of the exploits
  addPolylines(data = exploits_river, color = "#588BAD", weight = 1.5, options = pathOptions(pane = "river"), group = "exploits", opacity = 0.8) |>
  
## lakes of the exploits
  addPolygons(data = lakes, color = "#234667", stroke = FALSE, fillOpacity = 0.9, options = pathOptions(pane = "lakes"), group = "exploits") |>
  
  addScaleBar(position = "bottomleft") |> 
  addLogo(arrow, position = "bottomleft")



```


::: {.panel-tabset}

### Water Quality

Describe data here...   
  
Explore water quality variables from various tributaries across the Exploits watershed. Click on a site to see the data collected. The water quality layer selected is also plotted [below the map](#bottom5) against any other water quality variable you specify.
 
Link to download data is [below map](#bottom5).   

```{r}
#| echo: false
#| include: false

chem <- chem |> 
  clean_names() |> 
  mutate(x = as.numeric(x)*-1) |> 
  mutate(year_trib = paste0(year, "-", ifelse(trib_size == "Bg", "Big", "Small"))) |> 
  group_by(year, site_code, trib_size, year_trib, x, y) |> 
  summarise(across(.cols = c(water_temp, p_h, ec, dissolved_solids_ppm, alkalinity_ppm, doc, tdn, tdn_100), .fns = ~ mean(.x, na.rm = TRUE))) |>
  ungroup()
  
chem$year <- as.factor(chem$year)
chem$trib_size <- as.factor(chem$trib_size)
chem$year_trib <- as.factor(chem$year_trib)
chem$y <- as.numeric(chem$y)
chem$x <- as.numeric(chem$x)

names(chem)[names(chem) == 'water_temp'] <- 'Water temperature (˚c)'
names(chem)[names(chem) == 'p_h'] <- 'pH'
names(chem)[names(chem) == 'ec'] <- 'Electrical conductivity (µS/cm)'
names(chem)[names(chem) == 'dissolved_solids_ppm'] <- 'Dissolved solids (ppm)'
names(chem)[names(chem) == 'alkalinity_ppm'] <- 'Alkalinity (ppm)'
names(chem)[names(chem) == 'doc'] <- 'Dissolved organic carbon (mg/L)'
names(chem)[names(chem) == 'tdn'] <- 'Total dissolved nitrogen (mg/L)'



ojs_define(chem_data = chem)
```

```{r, palettes_chem}
#| include: false

pal.tp <- colorNumeric(palette = "inferno", domain = chem$`Water temperature (˚c)`)
pal.tp.leg <- colorNumeric(palette = "inferno", domain = chem$`Water temperature (˚c)`, reverse = T)

pal.ph <- colorNumeric(palette = "inferno", domain = chem$`pH`)
pal.ph.leg <- colorNumeric(palette = "inferno", domain = chem$`pH`, reverse = T)

pal.ec <- colorNumeric(palette = "inferno", domain = chem$`Electrical conductivity (µS/cm)`)
pal.ec.leg <- colorNumeric(palette = "inferno", domain = chem$`Electrical conductivity (µS/cm)`, reverse = T)

pal.ds <- colorNumeric(palette = "inferno", domain = chem$`Dissolved solids (ppm)`)
pal.ds.leg <- colorNumeric(palette = "inferno", domain = chem$`Dissolved solids (ppm)`, reverse = T)

pal.ak <- colorNumeric(palette = "inferno", domain = chem$`Alkalinity (ppm)`)
pal.ak.leg <- colorNumeric(palette = "inferno", domain = chem$`Alkalinity (ppm)`, reverse = T)

pal.dc <- colorNumeric(palette = "inferno", domain = chem$`Dissolved organic carbon (mg/L)`)
pal.dc.leg <- colorNumeric(palette = "inferno", domain = chem$`Dissolved organic carbon (mg/L)`, reverse = T)

pal.dn <- colorNumeric(palette = "inferno", domain = chem$`Total dissolved nitrogen (mg/L)`)
pal.dn.leg <- colorNumeric(palette = "inferno", domain = chem$`Total dissolved nitrogen (mg/L)`, reverse = T)


```

```{r, icons}
#| include: false
icon_tp <- awesomeIcons(
  icon = c("circle", bs_icon("suit-diamond-fill"), bs_icon("square-fill"))[match(chem$year_trib, c("2024-Big", "2024-Small", "2025-Small"))], # Map factor levels to icon names
  library = "fa",
  iconColor = pal.tp(chem$`Water temperature (˚c)`))

icon_ph <- awesomeIcons(
  icon = c("circle", bs_icon("suit-diamond-fill"), bs_icon("square-fill"))[match(chem$year_trib, c("2024-Big", "2024-Small", "2025-Small"))], # Map factor levels to icon names
  library = "fa",
  iconColor = pal.ph(chem$`pH`))


icon_ec <- awesomeIcons(
  icon = c("circle", bs_icon("suit-diamond-fill"), bs_icon("square-fill"))[match(chem$year_trib, c("2024-Big", "2024-Small", "2025-Small"))], # Map factor levels to icon names
  library = "fa",
  iconColor = pal.ec(chem$`Electrical conductivity (µS/cm)`))


icon_ds <- awesomeIcons(
  icon = c("circle", bs_icon("suit-diamond-fill"), bs_icon("square-fill"))[match(chem$year_trib, c("2024-Big", "2024-Small", "2025-Small"))], # Map factor levels to icon names
  library = "fa",
  iconColor = pal.ds(chem$`Dissolved solids (ppm)`))


icon_ak <- awesomeIcons(
  icon = c("circle", bs_icon("suit-diamond-fill"), bs_icon("square-fill"))[match(chem$year_trib, c("2024-Big", "2024-Small", "2025-Small"))], # Map factor levels to icon names
  library = "fa",
  iconColor = pal.ak(chem$`Alkalinity (ppm)`))


icon_dc <- awesomeIcons(
  icon = c("circle", bs_icon("suit-diamond-fill"), bs_icon("square-fill"))[match(chem$year_trib, c("2024-Big", "2024-Small", "2025-Small"))], # Map factor levels to icon names
  library = "fa",
  iconColor = pal.dc(chem$`Dissolved organic carbon (mg/L)`))


icon_dn <- awesomeIcons(
  icon = c("circle", bs_icon("suit-diamond-fill"), bs_icon("square-fill"))[match(chem$year_trib, c("2024-Big", "2024-Small", "2025-Small"))], # Map factor levels to icon names
  library = "fa",
  iconColor = pal.dn(chem$`Total dissolved nitrogen (mg/L)`),
  extraClasses = "down-6")



iconSet <- awesomeIconList(
  "2024-Big" = makeAwesomeIcon(
    icon = "circle",
    library = 'fa'
  ),
  "2024-Small" = makeAwesomeIcon(
    icon = bs_icon("suit-diamond-fill")
  ),
  "2025-Small" = makeAwesomeIcon(
    icon = bs_icon("square-fill")
  )
)


```

```{r, popup_labels_chem}
#| include: false
# adds information to popups for each reach when it is selected


label_text_chem <- glue("<b><u>Tributary:</u></b> <br/>",
  "<b>Water Temperature: </b> {round(chem$`Water temperature (˚c)`, 1)} ˚c<br/>",                      
  "<b>pH: </b> {round(chem$pH, 2)}<br/>",
  "<b>Electrical conductivity:</b> {round(chem$`Electrical conductivity (µS/cm)`,1)} µS/cm<br/>",
  "<b>Dissolved Solids: </b> {round(chem$`Dissolved solids (ppm)`,1)} ppm<br/>",
  "<b>Alkalinity: </b> {round(chem$`Alkalinity (ppm)`,1)} ppm<br/>",
  "<b>Dissolved organic carbon: </b> {round(chem$`Dissolved organic carbon (mg/L)`,2)} mg/L<br/>",
  "<b>Total dissolved nitrogen: </b> {round(chem$`Total dissolved nitrogen (mg/L)`,2)} mg/L<br/>") |> 
 
  lapply(htmltools::HTML)
```

```{r, map_chem}
#| echo: false
#| column: screen 
#| warning: false


exploits_map_chem <- exploits_map |> 
  
  addMapPane("data_tp", zIndex = 455) |> 
  addMapPane("data_ph", zIndex = 450) |> 
  addMapPane("data_ec", zIndex = 445) |> 
  addMapPane("data_ds", zIndex = 440) |> 
  addMapPane("data_ak", zIndex = 435) |> 
  addMapPane("data_dc", zIndex = 430) |> 
  addMapPane("data_dn", zIndex = 425) |> 

## temp
  addAwesomeMarkers(data = chem, lng = ~ chem$x, lat = ~chem$y, options = pathOptions(pane = "data_tp"), group = "Water temperature (˚c)", layerId = ~`Water temperature (˚c)`, icon = icon_tp, popup = ~ label_text_chem) |>
      
  addLegend(pal = pal.tp.leg, values = chem$`Water temperature (˚c)`, title = "Water temperature (˚c)", position = 'topright', group = "Water temperature (˚c)", opacity = 1, na.label = "NA", labFormat = labelFormat(transform = function(x) sort(x, decreasing = TRUE))) |>
  
## pH
  addAwesomeMarkers(data = chem, lng = ~ chem$x, lat = ~chem$y, options = pathOptions(pane = "data_ph"), group = "pH", layerId = ~pH, icon = icon_ph, popup = ~ label_text_chem) |>
      
  addLegend(pal = pal.ph.leg, values = chem$`pH`, title = "pH", position = 'topright', group = "pH", opacity = 1, na.label = "NA", labFormat = labelFormat(transform = function(x) sort(x, decreasing = TRUE))) |>
  
  addLegendAwesomeIcon(iconSet = iconSet,
                       title = "Year - Tributary Size",
                       position = 'topright',
                       group = 'year-trib')  |>

## ec  
  addAwesomeMarkers(data = chem, lng = ~ chem$x, lat = ~chem$y, options = pathOptions(pane = "data_ec"), group = "Electrical conductivity (µS/cm)", layerId = ~`Electrical conductivity (µS/cm)`, icon = icon_ec, popup = ~ label_text_chem) |>
      
  addLegend(pal = pal.ec.leg, values = chem$`Electrical conductivity (µS/cm)`, title = "Electrical conductivity (µS/cm)", position = 'topright', group = "Electrical conductivity (µS/cm)", opacity = 1, na.label = "NA", labFormat = labelFormat(transform = function(x) sort(x, decreasing = TRUE))) |>  
  
## ds 
  addAwesomeMarkers(data = chem, lng = ~ chem$x, lat = ~chem$y, options = pathOptions(pane = "data_ds"), group = "Dissolved solids (ppm)", layerId = ~ `Dissolved solids (ppm)`, icon = icon_ds, popup = ~ label_text_chem) |>
      
  addLegend(pal = pal.ds.leg, values = chem$`Dissolved solids (ppm)`, title = "Dissolved Solids (ppm)", position = 'topright', group = "Dissolved solids (ppm)", opacity = 1, na.label = "NA", labFormat = labelFormat(transform = function(x) sort(x, decreasing = TRUE)))  |>

## ak 
  addAwesomeMarkers(data = chem, lng = ~ chem$x, lat = ~chem$y, options = pathOptions(pane = "data_ak"), group = "Alkalinity (ppm)", layerId = ~ `Alkalinity (ppm)`, icon = icon_ak, popup = ~ label_text_chem) |>
      
  addLegend(pal = pal.ak.leg, values = chem$`Alkalinity (ppm)`, title = "Alkalinity (ppm)", position = 'topright', group = "Alkalinity (ppm)", opacity = 1, na.label = "NA", labFormat = labelFormat(transform = function(x) sort(x, decreasing = TRUE)))  |>
  
## dc 
  addAwesomeMarkers(data = chem, lng = ~ chem$x, lat = ~chem$y, options = pathOptions(pane = "data_dc"), group = "Dissolved organic carbon (mg/L)", layerId = ~ `Dissolved organic carbon (mg/L)`, icon = icon_dc, popup = ~ label_text_chem) |>
      
  addLegend(pal = pal.dc.leg, values = chem$`Dissolved organic carbon (mg/L)`, title = "Dissolved organic carbon (mg/L)", position = 'topright', group = "Dissolved organic carbon (mg/L)", opacity = 1, na.label = "NA", labFormat = labelFormat(transform = function(x) sort(x, decreasing = TRUE)))  |>
  
## dn
  addAwesomeMarkers(data = chem, lng = ~ chem$x, lat = ~chem$y, options = pathOptions(pane = "data_dn"), group = "Total dissolved nitrogen (mg/L)", layerId = ~ `Total dissolved nitrogen (mg/L)`, icon = icon_dn, popup = ~ label_text_chem) |>
      
  addLegend(pal = pal.dn.leg, values = chem$`Total dissolved nitrogen (mg/L)`, title = "TDN", position = 'topright', group = "Total dissolved nitrogen (mg/L)", opacity = 1, na.label = "NA", labFormat = labelFormat(transform = function(x) sort(x, decreasing = TRUE)))  |>
  
# ability to choose which layers to see
  addGroupedLayersControl(
    baseGroups = c("Satellite", "Topographic", "White & Grey"),
    overlayGroups = list(
      "Water quality data" = c("Water temperature (˚c)", "pH", "Electrical conductivity (µS/cm)", "Dissolved solids (ppm)", "Alkalinity (ppm)", "Dissolved organic carbon (mg/L)", "Total dissolved nitrogen (mg/L)")),
    position = "topleft",
    options = groupedLayersControlOptions(
      groupCheckboxes = FALSE,
      collapsed = FALSE,
      groupsCollapsable = FALSE,
      sortLayers = FALSE,
      sortGroups = FALSE,
      sortBaseLayers = FALSE,
      exclusiveGroups = c("Water quality data"),
    )
  )  |> 

hideGroup(c("pH", "Electrical conductivity (µS/cm)", "Dissolved solids (ppm)", "Alkalinity (ppm)", "Dissolved organic carbon (mg/L)", "Total dissolved nitrogen (mg/L)"))  |> 

htmlwidgets::onRender("
    function(el, x) {
      this.on('overlayadd', function(e) {
        // Dispatch custom event with the layer name
        window.dispatchEvent(new CustomEvent('leafletLayerSelected', {
          detail: e.name
        }));
      });
    }
  ")
  

exploits_map_chem

```

\  

```{ojs}
//| echo: false
//| output: false

import {legend} from "@d3/color-legend"

htmlToImage = import("https://cdn.skypack.dev/html-to-image@1.11.1?min")


image = async (node) => {
  const tabPanel = node.closest('.tab-pane').classList.add('active', 'show');
  
  let originalDisplay = '';
  if (tabPanel && tabPanel.style.display === 'none') {
    originalDisplay = 'none';
    tabPanel.style.display = 'block';
  }
  
  const n = d3.select(node);
  const pb = n.style("padding-bottom");
  const bg = n.style("background-color");
  
  if (node.nodeName === "FIGURE") {
    n.style("padding-bottom", ".4em");
    n.style("background-color", "white");
  }
  const im = await htmlToImage.toCanvas(node, { pixelRatio: 2 });
  im.style = `width: ${im.width >> 1}px`;
  if (pb) n.style("padding-bottom", pb); // restores the original styles.
  if (bg) n.style("background-color", bg);
  return im;
}
```

```{ojs}
//| echo: false
//| panel: fill

viewof selectedLayerGroup = {
  // 1. Create a container for the entire viewof component
  let container = document.createElement("div");
  
  // 2. Create the fixed label and the dynamic display area
  container.innerHTML = `
    <div style="font-weight: bold; margin-bottom: 5px;">Map layer as X Axis:</div>
    <div id="selection-display">Water temperature (˚c)</div>
  `;
  
  // 3. Define which element actually holds the value for Observable
  const displayArea = container.querySelector("#selection-display");
  container.value = "Water temperature (˚c)";

  // 4. Listen for the custom event
  window.addEventListener("leafletLayerSelected", (e) => {
    // Update only the dynamic area
    displayArea.innerHTML = e.detail;
    
    // Update the value property of the container
    container.value = e.detail;
    
    // Dispatch input event to update Observable
    container.dispatchEvent(new Event("input", {bubbles: true}));
  });

  return container;
}


viewof yVariable = Inputs.select(["Water temperature (˚c)", "pH", "Electrical conductivity (µS/cm)", "Dissolved solids (ppm)", "Alkalinity (ppm)", "Dissolved organic carbon (mg/L)", "Total dissolved nitrogen (mg/L)"], {value: "Electrical conductivity (µS/cm)", label: html`<b>Select Y Axis:</b>`})


myPlot = Plot.plot({
    grid: true,
    color: {scheme: "inferno",
      legend: false},
    symbol: {legend: true, 
    range: ["circle", "diamond", "square"]
    },
    marks: [
      Plot.dot(data, {
        x: selectedLayerGroup, 
        y: yVariable,
        fill: selectedLayerGroup,
        symbol: "year_trib"})
    ]
  })

data = transpose(chem_data)
```

```{ojs}
//| echo: false
//| output: false

plotImage = image(myPlot)


```


```{ojs}
//| echo: false

viewof downloadSVG = Inputs.button("Download Plot PNG", {
  reduce: () => {
    var canvas = document.querySelector('canvas');
    var image = canvas.toDataURL("image/png").replace("image/png", "image/octet-stream");
  
    var element = document.createElement('a');
    var filename = 'water_quality_plot.png';
    element.setAttribute('href', image);
    element.setAttribute('download', filename);

    element.click();
  }
})
```

\   
Describe water chem data and data collection...    
What is a big trib vs a small trib...  
  
[Click here to download the Exploits River water quality data](https://github.com/emmwilson/NL_watersheds/blob/0d9a7a8e31f9d6fc5a61aac211dbf85ce949138c/data/spatial%20data/Water_Quality_data_explore.xlsx?raw=true)   
  
Citation: ...  
  
<div id="bottom5"></div>   




### Invertebrates and periphyton
  
Explore the invertebrate an



:::