---
title: "`r params$watershed` watershed predictions"
theme: sandstone
params:
    watershed: "Gander"
format:
  html:
    css: styles.css
    grid:
      sidebar-width: 0px
      margin-width: 0px
      body-width: 2500px
engine: knitr
---


HydroATLAS data that can be used in gravel models to estimate gravel with the proper size for salmon spawning. Includes example predictions of gravel size.
Click on a reach to see the data for each segment. Description of map contents and a link to download  map shapefile are [below map](#bottom).

::: {.column-screen}
```{r, library}
#| include: false
pacman::p_load(shiny,
               bslib,
               ggplot2,
               ggtext,
               shinythemes,
               thematic,
               tidyverse,
               httr,
               leaflet,
               leafem,
               terra,
               htmltools,
               glue,
               cowplot,
               grDevices,
               leaflegend,
               smoothr,
               htmlwidgets)

if (!requireNamespace("leaflet.extras", quietly = TRUE)) {
  remotes::install_github("trafficonese/leaflet.extras")
}
library(leaflet.extras)

knitr::read_chunk('addGroupedLayersControl.R')

```

```{r, import}
#| include: false
# import layers and ensure all in same projection
## river watershed
file_watershed <- paste0("data/spatial data/watersheds/watershed_", params$watershed, ".shp")

river_drainage <- vect(file_watershed) |> 
  project("EPSG:4326 - WGS 84") 
  
bbox <- ext(river_drainage)
  
  
  ## stream and rivers in watershed
file_river <- paste0("data/spatial data/watersheds/river_", params$watershed, ".shp")

rivers <- vect(file_river) |>  
    project("EPSG:4326 - WGS 84")
  
  ## lakes
file_lakes <- paste0("data/spatial data/watersheds/lakes_", params$watershed, ".shp")

lakes <- vect(file_lakes) |> 
    project("EPSG:4326 - WGS 84")
  
## bounding rectangle for NL
NL <- vect("data/spatial data/NL_boundaries.shp") |>  
  project("EPSG:4326 - WGS 84")  

## areas blocked by CABD 
file_barriers <- paste0("data/spatial data/watersheds/barriers_", params$watershed, ".shp")
barriers <- vect(file_barriers)

file_inaccess <- paste0("data/spatial data/watersheds/inaccess_", params$watershed, ".shp")
inaccess <- vect(file_inaccess)


  # import arrow
arrow <- "https://cdn.pixabay.com/photo/2013/07/12/17/54/arrow-152596_960_720.png"
```

```{r, predictions}
#| include: false
#| warning: false
#| echo: false

# use river reach data to predict outcomes

# values from Ky's presentation
ps <- 2650 # kg/m3, density of sediment
p <- 2000 # kg/m3, density of water
g <- 9.80665 # m/s2, acceleration by gravity
n <- 0.04 # channel roughness coefficient
t <- 0.04 # shields parameter

## predict responses based on data from field at each reach
rivers$width <- (rivers$ria_ha_csu/rivers$LENGTH_KM)*10

rivers$tb <- p*g*(n^(3/5))*((rivers$DIS_AV_CMS/rivers$width)^(3/5))*((rivers$sgr_dk_rav/10000)^(7/10))  

rivers$D50 <- (rivers$tb/((ps-p)*g*t))*1000 


```

```{r, palettes}
#| include: false
#| warning: false
#| echo: false
 #colour palettes for predictions
  # invert.pal <- colorNumeric(c("#f47c3c", "#d9534f"), domain = predictions_reaches$invert, na.color = "transparent")
  # embed.pal <- colorNumeric(c("#e9c602", "#93c54b"), domain =  predictions_reaches$embed, na.color = "transparent")
top_discharge <- quantile(rivers$DIS_AV_CMS, 0.9, na.rm = T)

discharge.pal <- colorNumeric(c("#dfdce3", "#600075"), domain = c(0, top_discharge), na.color = "transparent")
discharge.col <- discharge.pal(rivers$DIS_AV_CMS)

discharge.pal90 <- colorNumeric(c("#600075"), domain = c(top_discharge, max(rivers$DIS_AV_CMS)), na.color = "transparent") 
discharge.col90 <- discharge.pal90(rivers$DIS_AV_CMS)

# colour palette for legend so high numbers at top
discharge.col.leg <- discharge.pal(seq(0, top_discharge, by = top_discharge/7))
discharge.pal.leg <- colorNumeric(rev(discharge.col.leg), domain = c(0, top_discharge), na.color = "transparent")  


# colour palette for width
# make top 90th percentile one colour
top_width <- quantile(rivers$width, 0.9, na.rm = T)

width.pal <- colorNumeric(c("#F5DAEA", "#B50061"), domain = c(0, top_width), na.color = "transparent") # 
width.col <- width.pal(rivers$width)

width.pal90 <- colorNumeric(colorRamp(c("#B50061")), domain = c(top_width, max(rivers$width)), na.color = "transparent") 
width.col90 <- width.pal90(rivers$width)
  
# colour palette for legend so high numbers at top
width.col.leg <- width.pal(seq(0, top_width, by = top_width/7))
width.pal.leg <- colorNumeric(rev(width.col.leg), domain = c(0, top_width), na.color = "transparent")  

#width.pal <- colorFactor(c("#f2ca97", "#d4021d"), domain = exploits_river$width*1000, na.color = "transparent") use this function for any categorical data you want to include in the map
  
# colour palette for stream gradient

# make top 90th percentile one colour
top_gradient <- quantile(rivers$sgr_dk_rav/10000, 0.9, na.rm = T)

gradient.pal <- colorNumeric(colorRamp(c("white", "#1C1C1C")), domain = c(0, top_gradient), na.color = "transparent") 
gradient.col <- gradient.pal(rivers$sgr_dk_rav/10000)

gradient.pal90 <- colorNumeric(colorRamp(c("#1C1C1C")), domain = c(top_gradient, max(rivers$sgr_dk_rav/10000)), na.color = "transparent") 
gradient.col90 <- gradient.pal90(rivers$sgr_dk_rav/10000)


# colour palette for legend so high numbers at top
gradient.col.leg <- gradient.pal(seq(0, top_gradient, by = top_gradient/7))
gradient.pal.leg <- colorNumeric(rev(gradient.col.leg), domain = c(0, top_gradient), na.color = "transparent")  

# colour palette for D50, with 16-64 green and over 90 orange
  
D50.pal <- colorNumeric(c("#e9c602","#e9c602", "#e9c602","#FD602F"), domain = c(0, 16, 64, ceiling(max(rivers$D50, na.rm = T)/16)*16), na.color = "brown")
 
 # if D50s reach over 200 then use 200 as cap, but if not then use closest multiple of 10 to max D50 as cap
top_D50 <- ifelse(max(rivers$D50, na.rm = T) >= 200, 200, ceiling(max(rivers$D50, na.rm = T)/10)*10)

 
 # map colours
D50.col <-  D50.pal(rivers$D50)
D50.col[rivers$D50 >= 16 & rivers$D50 <= 64] <- "#00D458"
D50.col[rivers$D50 >= top_D50] <- "#FD602F" 

  #legend
D50.col.leg <- D50.pal(seq(0, top_D50, by = top_D50/10))
D50.col.leg[seq(0, top_D50, by = top_D50/10) >= 16 & seq(0, top_D50, by = top_D50/10) <= 64] <- "#00D458"
D50.col.leg[seq(0, top_D50, by = top_D50/10) >= top_D50] <- "#FD602F"

D50.pal.leg <- colorNumeric(rev(D50.col.leg), domain = c(0,top_D50), na.color = "brown")

```


```{r, popup_labels}
#| include: false
# adds information to popups for each reach wehn it is selected

  label_text <- glue("<b><u>{'Segment info:'}</u></b> <br/>",
"<b>Predicted gravel size (D50): </b> {round(rivers$D50, 1)} mm<br/>",
"<b>Discharge: </b> {round(rivers$DIS_AV_CMS, 1)} m<sup>3</sup>/s<br/>",
"<b>Width: </b> {round(rivers$width,1)} m<br/>",
"<b>Gradient: </b> {round(rivers$sgr_dk_rav/10,1)} mm/m<br/>") |> 

  lapply(htmltools::HTML)
```

```{r, map}
#| echo: false
#| warning: false


map <- leaflet(width = "100vw", height = "75vh", options = leafletOptions(zoomSnap = 0.25)) |> 
  # registerPlugin(groupedLayerControlPlugin) |> 

  addProviderTiles('Esri.WorldImagery', group = "Satellite") |> # add base map options
  addProviderTiles('Esri.WorldTopoMap', group = "Topographic") |> # adds base map
  addProviderTiles('Esri.WorldGrayCanvas', group = "White & Grey") |> 

      # open map to this screen
      fitBounds(as.numeric(bbox[1]), as.numeric(bbox[3]), as.numeric(bbox[2]), as.numeric(bbox[4])) |> 

      #can recenter to that location by clicking button
      addHomeButton(ext = c(as.numeric(bbox[1]), as.numeric(bbox[3]), as.numeric(bbox[2]), as.numeric(bbox[4])), position = "topleft", group = "Recenter") |> 

      # create layers to add vectors to
      addMapPane("NL", zIndex = 400) |> 
      addMapPane("drainage", zIndex = 410) |> 
      addMapPane("rivers", zIndex = 415) |> 
      addMapPane("lakes", zIndex = 420) |>
      addMapPane("data_D50", zIndex = 450) |> 
      addMapPane("data_disch", zIndex = 445) |> 
      addMapPane("data_disch90", zIndex = 444) |> 
      addMapPane("data_width", zIndex = 440) |> 
      addMapPane("data_width90", zIndex = 439) |> 
      addMapPane("data_grad", zIndex = 435) |>
      addMapPane("data_grad90", zIndex = 434) |>
      addMapPane("inaccess", zIndex = 455) |> 
      addMapPane("barriers", zIndex = 460) |> 
  
      
      #addPolygons(data = NL, color = "#82797F", weight = 1.5, fillColor = "#82797F", fillOpacity = 0.5, options = pathOptions(pane = "NL")) |> ## need a new NL file
      # add polygons and lines
      ## exploits drainage area
      addPolygons(data = river_drainage, color = "#32B1DB", weight = 1.25, fillColor = "#236D9E", opacity = 0.4, fillOpacity = 0.1, options = pathOptions(pane = "drainage")) |>
  
      ## rivers of the rocky
      addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = "#588BAD", weight = rivers$width^(1/3), options = pathOptions(pane = "rivers"), group = "exploits", opacity = 0.8) |>

  ## lakes of the exploits
      addPolygons(data = lakes, color = "#234667", stroke = FALSE, fillOpacity = 0.9, options = pathOptions(pane = "lakes"), group = "exploits") |>

  ## add barriers  
  addPolylines(data = smooth(inaccess, method = "ksmooth", smoothness = 0.5), color = "#C41F02", weight = 2.5, options = pathOptions(pane = "inaccess"), group = "CABD barriers and areas past", opacity = 1, popup = ~label_text) |>
  
  addCircleMarkers(data = barriers, color = "#720F0F", weight = 1, fillColor = "#720F0F", radius = 3.5, fillOpacity = 1, fill = T, opacity = 1, options = pathOptions(pane = "inaccess"), group = "CABD barriers and areas past", popup = ~label_text) |> 
  
  addLegend(colors = "#C41F02", opacity = 1, labels = "<b>CABD barriers and areas past</b>", group = "CABD barriers and areas past", position = "topright") |> 
  
  ### D50
      addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = D50.col, weight = rivers$width^(1/3), fillOpacity = 1, opacity = 1, options = pathOptions(pane = "data_D50"), group = "D50", popup = ~label_text) |>
      
      addLegend(pal = D50.pal.leg, values = seq(0, top_D50, by = top_D50/10), title = "D50 gravel size (mm)", position = 'topright', group = "D50", opacity = 1, labFormat = function(type, cuts, p) {
      paste0(rev(c(as.character(seq(0, top_D50-top_D50/10, by = top_D50/10)), ifelse(top_D50 >= 200, "≥200", paste0('≥ ', top_D50))))) 
    }) |>  
  
  
      ### discharge
      addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = ~discharge.pal(rivers$DIS_AV_CMS), weight = rivers$width^(1/3)-0.05, fillOpacity = 1, opacity = 1, options = pathOptions(pane = "data_disch"), group = "Discharge", popup = ~label_text) |>
       
      addLegend(pal = discharge.pal.leg, values = seq(0, top_discharge, by = top_discharge/7), title = "Stream discharge (m<sup>3</sup>/s)", position = 'topright', group = "Discharge", opacity = 1, labFormat = function(type, cuts, p) {
      paste0(rev(c(as.character(round(seq(0, top_discharge-top_discharge/7, by = top_discharge/7), 1)), paste0('≥ ', round(top_discharge, 1)))))
        }) |>
   
       addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = ~discharge.pal90(rivers$DIS_AV_CMS), weight = rivers$width^(1/3)-0.05, fillOpacity = 1, opacity = 1, options = pathOptions(pane = "data_disch90"), group = "Discharge", popup = ~label_text)  |># add in values above top quantile
  
      ### width
      addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = ~width.pal(rivers$width), weight = rivers$width^(1/3)-0.1, fillOpacity = 1, opacity = 1, options = pathOptions(pane = "data_width"), group = "Width", popup = ~label_text) |>

      addLegend(pal = width.pal.leg, values = seq(0, top_width, by = top_width/7), title = "Stream width (m)", position = 'topright', group = "Width", opacity = 1, labFormat = function(type, cuts, p) {
      paste0(rev(c(as.character(round(seq(0, top_width-top_width/7, by = top_width/7),1)), paste0('≥ ', round(top_width, 1)))))
        }) |> 
      
      addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = ~width.pal90(rivers$width), weight = rivers$width^(1/3)-0.1, fillOpacity = 1, opacity = 1, options = pathOptions(pane = "data_width90"), group = "Width", popup = ~label_text)  |># add in values above top quantile

  
      ### gradient
      addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = ~gradient.pal(rivers$sgr_dk_rav/10000), weight = rivers$width^(1/3)-0.15,  opacity = 1, options = pathOptions(pane = "data_grad"), group = "Gradient", popup = ~label_text) |>
      
      addLegend(pal = gradient.pal.leg, values = seq(0, top_gradient, by = top_gradient/7), title = "Stream gradient (mm/m)", position = 'topright', group = "Gradient", opacity = 1, labFormat = function(type, cuts, p) {
      paste0(rev(c(as.character(round(seq(0, (top_gradient*1000)-(top_gradient*1000)/7, by = (top_gradient*1000)/7), 1)), paste0('≥ ', round(top_gradient*1000, 1)))))
        }) |> 
      
  addPolylines(data = smooth(rivers, method = "ksmooth", smoothness = 0.5), color = ~gradient.pal90(rivers$sgr_dk_rav/10000), weight = rivers$width^(1/3)-0.15, opacity = 1, options = pathOptions(pane = "data_grad90"), group = "Gradient", popup = ~label_text)  |># add in values above top quantile
  
      # ability to choose which layers to see
  addGroupedLayersControl(
    baseGroups = c("Satellite", "Topographic", "White & Grey"),
    overlayGroups = list(
      "Predictions and HydroSHEDs data" = c("D50", "Discharge", "Width", "Gradient"),
      "Barriers" = c("CABD barriers and areas past")
    ),
    position = "topleft",
    options = groupedLayersControlOptions(
      groupCheckboxes = FALSE,
      collapsed = FALSE,
      groupsCollapsable = FALSE,
      sortLayers = FALSE,
      sortGroups = FALSE,
      sortBaseLayers = FALSE,
    )
  ) |> 
  
  hideGroup("CABD barriers and areas past") |> 
  
  addScaleBar(position = "bottomleft") |> 
  
  addLogo(arrow, position = "bottomleft", offset.x = 2, offset.y = 40)
  

map
```
<div id="bottom"></div>
|  
| <u><b>Description of map contents:</b></u> 
|  
| <b>D50 Gravel size:</b>
|  
| Gravel size is important to the Atlantic Salmon spawning; too fine or too coarse, and the salmon will either not spawn or have poor egg survival (Purchase 2018). We considered 16 - 64 mm to be the most suitable substrate size, and this is represented as green on th D50 map. However, salmon may spawn in substrate that isn't ideal, represented by yellow on the map. Finally, unsuitable substrate is represented by orange.
| 
| D50 was predicted based on the following calculations (Wilkins & Snyder 2011)  
|  
| D50 = 𝛕<sub>b</sub> / (𝛒<sub>s</sub> - 𝛒) * g * 𝛕<sup>*</sup>   
|  
| <u>Parameters:</u>   
| 𝛕<sub>b</sub> = 𝛒 * g * n<sup>3/5</sup> * [Q/W]<sup>3/5</sup> * S<sup>7/10</sup>  
| 𝛒 = density of water = 1000 kg/m<sup>3</sup>  
| g = acceleration by gravity = 9.80665 m/s<sup>2</sup>  
| n = channel roughness coefficient = n = 0.04 as estimate for gravel bed rivers from Barnes 1967  
| 𝛒<sub>s</sub> = sediment density = 2650 kg/m<sup>3</sup>  
| 𝛕<sup>*</sup>  =  Shields parameter = 0.04 to follow Wilkins & Snyder 2011  
|  
| <u>Variables to retreive from HydroSHEDs:</u>  
| S = channel gradient, Calculated as the ratio between the elevation drop within the river reach (i.e. the difference between min. and max. elevation along the reach) and the length of the reach.  
| Q = Discharge (m<sup>3</sup>/s), Average long-term discharge estimate for river reach, in cubic meters per second.    
| W = channel width (m), Calculated as the surface river area of the river reach segment divided by length of the river reach segment.    
|  
| <b>CABD barriers and areas past:</b>
| Potential barriers to upstream salmon movement were taken from the Canadian Aquatic Barriers Database. Barriers were filtered to remove barriers that are insufficient to impede movement: a fishway exists, barrier has been decommissioned/removed, barrier was a dam less than 5 m high, or barrier's Used for Network Analysis field was false. Any stream segment upstream of the barrier is considered potentially inaccessible. Many of these barriers may in fact be passable by salmon (e.g. passable rapid listed as waterfall) but we lack sufficient information to remove them from list the of potential barriers. Therefore, each barrier must be assessed on an individual basis if being incorporated into management or research plans.    
|  
| <b> Data and citations: </b>
| [Click here to download the `r params$watershed` HydroATLAS data shapefile](`r paste0("https://github.com/emmwilson/NL_watersheds/blob/0d9a7a8e31f9d6fc5a61aac211dbf85ce949138c/data/spatial%20data/watersheds/river_",params$watershed,".shp?raw=true")`)  
| 
| Barnes, H. 1967. Roughness Characteristics of Natural Channels. United States Geological Survey: Publications.  
| Canadian Wildlife Federation. (2024). Canadian Aquatic Barriers Database (CABD). Retrieved from https://aquaticbarriers.ca/  
| Linke, S., Lehner, B., Ouellet Dallaire, C., Ariwi, J., Grill, G., Anand, M., Beames, P., Burchard-Levine, V., Maxwell, S., Moidu, H., Tan, F., Thieme, M. (2019). Global hydro-environmental sub-basin and river reach characteristics at high spatial resolution. Scientific Data 6: 283. doi: https://doi.org/10.1038/s41597-019-0300-6   
| Lehner, B., Messager, M.L., Korver, M.C., Linke, S. (2022). Global hydro-environmental lake characteristics at high spatial resolution. Scientific Data 9: 351. doi: https://doi.org/10.1038/s41597-022-01425-z   
| Purchase, C. (2018). A systematic review on the effectiveness of salmonid spawning habitat improvements, and recommendations to potentially increase productivity of depressed Newfoundland Atlantic salmon (Salmo salar) populations.  
| Wilkins, and Snyder. 2011. Geomorphic comparison of two Atlantic coastal rivers: Toward an understanding of physical controls on Atlantic salmon habitat. River Research and Applications.  
:::